### 2016-05-10 ch01-intro/introduction.tex
Some diction suggestions

\section{Modeling}
Models play an important role in today's world, even though we often are not always aware of them.
In particular, models enable us to predict the future (within a certain bound of uncertainty):
\begin{itemize}
  \item supermarkets try to predict how much fruit and vegetables their regular customers buy such that they can order a large enough supply,
  \item traders on the stock market aim to predict whether companies such as Tesla or Apple will continue to prosper over the next quarter such that they can decide if it's time to buy more stock or rather sell,
  \item weather forecasters hope to predict what weather it might become tomorrow,
  \item how much electricity will the whole country use next week (do we need to start up a new power plant or not),
  \item how much noise the bodywork of a car will produce when it is driving over a bumpy road,
  \item what the exhaust of nitrogen oxides of a car will be for a certain operating point of a diesel engine, 
  \item are the vibrations in a robot are small enough for it to perform very precise tasks such as positioning electronic components with sub-millimeter precision,
\end{itemize}
and so on.

@@

\paragraph{Experiment design}
In the first step, an experiment is constructed to gain insight into the behavior of a system.
For system identification, experiment design is a question of designing input signals that will be applied to the system under test.
Typically, the design consists of constructing a signal such that the obtained model  has a small uncertainty in some sense~\citep{Goodwin1977,Mehra1974,Goodwin2006GBO,Levadi1966}.
On the one hand, if one has very precise prior information, one could even construct optimal signals~\citep{Gagliardi1967,Karlin1966,Gevers2011ExpDesign,Zarrop1979}.
Typically, the design is very sensitive to the prior knowledge of the system.
This is necessary to focus all the energy in the signal towards exciting the system dynamics as to reduce the uncertainty in the model.
However, when the prior knowledge is incomplete, uncertain or even incorrect, optimal signals can lead to very poor models as they might not excite other dynamics.
An alternative approach is to incorporate the uncertainty of the prior knowledge into the signal design.
This leads to so-called robust optimal experiment design~\citep{Rojas2012,Goodwin2006} that construct signals that optimize the worst-case model uncertainty.

@@

\paragraph{Select a model structure}
Once an experiment has been performed, the obtained input-output data is to be processed.
In many cases, the eventual goal of identification is to obtain a parametric model, so the user has to select a model structure and/or complexity based on both his prior knowledge and on the data.
A useful tool to help choose the model structure and/or model orders, is to inspect a nonparametric model of the system.
One such tool is the frequency response function.
For instance, by inspecting a frequency response function, a control engineer can have a rough idea of the system behavior in term of  approximate resonance frequencies, asymptotic slopes, \latin{etc.}
This gives a qualitative view of the minimally required complexity of the model and is hence a useful diagnostic tool.
However, since most parametric estimators enforce very little prior knowledge about the system, noise in the measurements can sometimes be a problem to distinguish the actual system features.

\paragraph{Estimate the model parameters}
Once the model structure has been chosen, its parameters are to be estimated.
This is often done by formulating an optimization problem which serves to minimize a particular cost function.
The cost function is in essence a measure of the differences between the measured data and the model evaluated for particular parameter values.
By minimizing the cost function, the optimization procedure yields parameter values that minimize the misfit between the data and the model.
Unfortunately, for many identification settings, these optimization problems are non-convex~\citep{Boyd2004}.
This means that initial values need to be available to start the optimization process.
Moreover, for poor starting values, optimization algorithms may get stuck in local minima of the cost function.
Consequently, without reasonable starting values, the `optimized' model may also be of poor quality.

@@

\section{User-Friendly System Identification}
In the previous section, it has been seen that some of the system identification steps require a significant amount of prior knowledge and/or craftsmanship of the user to make good choices.

>>>>

\section{Modeling}
Models play an important role in today's world, even though we often are not always aware of them.
In particular, models enable us to predict the future (within a certain bound of uncertainty):
\begin{itemize}
  \item supermarkets try to predict how much fruit and vegetables their regular customers buy such that they can order a large enough supply,
  \item traders on the stock market aim to predict whether companies such as Tesla or Apple will continue to prosper over the next quarter such that they can decide if it's time to buy more stock or rather sell,
  \item weather forecasters hope to predict what weather it might become tomorrow,
  \item how much electricity will the whole country use next week (do we need to start up a new power plant or not),
  \item how much noise the bodywork of a car will produce when it is driving over a bumpy road,
  \item what the exhaust of nitrogen oxides of a car will be for a certain operating point of a diesel engine, 
  \item are the vibrations in a robot are small enough for it to carry out very precise tasks such as positioning electronic components with sub-millimeter precision,
\end{itemize}
and so on.

@@

\paragraph{Experiment design}
In the first step, an experiment is constructed to gain insight into the behavior of a system.
For system identification, experiment design is a question of designing input signals that will be applied to the system under test.
Typically, the design consists of constructing a signal such that the obtained model  has a small uncertainty in some sense~\citep{Goodwin1977,Mehra1974,Goodwin2006GBO,Levadi1966}.
On the one hand, if one has very precise prior information, one could even construct optimal signals~\citep{Gagliardi1967,Karlin1966,Gevers2011ExpDesign,Zarrop1979}.
Typically, the design is very sensitive to the prior knowledge of the system.
This is necessary to focus all the energy in the signal towards exciting the system dynamics to the end of reducing the model uncertainty.
However, when the prior knowledge is incomplete, uncertain or even incorrect, optimal signals can lead to very poor models as they might not excite other dynamics.
An alternative approach is to incorporate the uncertainty of the prior knowledge into the signal design.
This leads to so-called robust optimal experiment design~\citep{Rojas2012,Goodwin2006} that construct signals that optimize the worst-case model uncertainty.

@@

\paragraph{Select a model structure}
Once an experiment has been performed, the obtained input-output data are to be processed.
In many cases, the eventual goal of identification is to obtain a parametric model, so the user has to select a model structure and complexity based on both his prior knowledge and on the data.
A useful tool to help choose the model structure and/or model orders, is to inspect a nonparametric model of the system.
One such tool is the frequency response function.
For instance, by inspecting a frequency response function, a control engineer can have a rough idea of the system behavior in term of  approximate resonance frequencies, asymptotic slopes, \latin{etc.}
This gives a qualitative view of the minimally required complexity of the model and is hence a useful diagnostic tool.
However, since most parametric estimators enforce very little prior knowledge about the system, noise in the measurements can sometimes be a problem to distinguish the actual system features.

\paragraph{Estimate the model parameters}
Once the model structure has been chosen, its parameters are to be estimated.
This is often done by formulating an optimization problem which serves to minimize a particular cost function.
The cost function is in essence a measure of the differences between the measured data and the model evaluated for particular parameter values.
By minimizing the cost function, the optimization procedure yields parameter values that minimize the misfit between the data and the model.
Unfortunately, for many identification settings, these optimization problems are non-convex~\citep{Boyd2004}.
This means that initial values need to be available to start the optimization process.
Moreover, for poor starting values, optimization algorithms can get stuck in local minima of the cost function.
Consequently, without reasonable starting values, the `optimized' model can also be of poor quality.

@@

\section{User-Friendly System Identification}
In the previous section, it has been seen that some of the system identification steps require a significant amount of prior knowledge and craftsmanship of the user to make good choices.

### 2016-05-10 ch03-lrm/nonparametric.tex
Remark: quality Hinf estimate
Some diction suggestions

Whereas many areas of system identification focus on parametric identification and/or parameter estimation, non-parametric techniques have received renewed interest from the community over the last decade.
One of the reasons for this is that non-parametric techniques typically require less assumptions and knowledge of the system than building parametric models.
As a consequence, non-parametric methods are often more robust than fully parametric methods.
Besides, \glspl{FRF} have been shown to provide a quick insight into the basic dynamic properties of \gls{LTI} systems.
This capability is very useful for system design, manual controller tuning, model validation and model selection for parametric models~\citep{Pintelon2012}.

@@

\paragraph*{Outline}
\secref{sec:theory} introduces theory of estimating \glspl{FRF} using local modeling techniques.
In \secref{sec:biascalc}, the bias of the \gls{LRM} is derived.
In \secref{sec:simulations} the performance of the different modeling methods are compared by means of simulations.
In \secref{sec:nparam:orderSelection}, some options for model order selection are discussed.
In \secref{sec:nonparametric:truncation}, an approach is developed to further reduce the variance of the \gls{LPM} estimate by truncation of the impulse response.
Finally, conclusions are presented in \secref{sec:conclusion}.

@@

\begin{remark}
This system can only be solved reliably if the columns in \eqref{eq:nparam:designMatrix} are linearly independent.
In practice, this corresponds to the requirements on the input spectrum as stated in \citet{Schoukens2009LPM}: the input spectrum should be sufficiently `rough' to separate the transient contribution from the contribution $\LocalModel{G} \LocalModel{U}$.
This can be obtained e.g. using random phase multisines or random input signals.
\end{remark}

@@

\section{Simulations}
\label{sec:simulations}

@@

\paragraph{Good signal-to-noise ratios}
The next situation we consider is when $\SNR=60\unit{dB}$ as shown in \figref{fig:nparam:comparison:hiSNR}.
Here, the \gls{LPM} remains severly biased for small datasets ($\nBW < 12$).
However, for sufficiently long experiments ($\nBW > 32$), the bias becomes negligible.
For the \gls{LRM} and \gls{LRIC}, the bias is always much smaller than the variance.
In contrast to \gls{LPM}, the \gls{LRM} already reaches a constant variance level when $\nBW > 10$.
It should also be noted that for the \gls{LRIC}, the variance is almost $10\unit{dB}$ more than the corresponding \gls{LPM} method.
It is hence not advantageous to use the \gls{LRIC} as-is.
For all methods (except the \gls{ETFE} and \gls{LRIC}), the simulations show that the \gls{RMSE} reaches the noise floor minus $\Delta \SNR$.

@@

\section{Conclusions}
\label{sec:conclusion}
In this chapter, the performance of non-parametric \gls{LPM} and \gls{LRM} estimators were compared to observe the \gls{FRF} of resonant systems.
Since the \gls{LRM} is a biased estimator while the \gls{LPM} is asymptotically unbiased, it is important to compare their behavior.
Hence, expressions for the bias of the \gls{LRM} were derived.
For good \glspl{SNR} in resonant systems, it was seen that the \gls{LRM} is always superior to the \gls{LPM}.
For extremely poor \gls{SNR}, i.e. $\SNR \approx 10\unit{dB}$, the \gls{LPM} may perform slightly better than the \gls{LRM} when short measurement records are available.
In this chapter, the \gls{LRIC} was introduced.
This is an extension of the \gls{LRM} which solves the non-linear optimization problem and should thus be able to avoid a bias altogether.
However, it is seen that this estimator is extremely noisy but seems to provide a smaller bias than the \gls{LRM}.
Since the overall \gls{RMS} error for the \gls{LRIC} is much larger than most other methods, the \gls{LRIC} cannot be used in its current form.

>>>>

Whereas many areas of system identification focus on parametric identification and/or parameter estimation, non-parametric techniques have received renewed interest from the community over the last decade.
One of the reasons for this is that non-parametric techniques typically require fewer assumptions and knowledge of the system than building parametric models.
As a consequence, non-parametric methods are often more robust than fully parametric methods.
Besides, \glspl{FRF} have been shown to provide a quick insight into the basic dynamic properties of \gls{LTI} systems.
This capability is very useful for system design, manual controller tuning, model validation and model selection for parametric models~\citep{Pintelon2012}.

@@

\paragraph*{Outline}
\secref{sec:theory} introduces theory of estimating \glspl{FRF} using local modeling techniques.
In \secref{sec:biascalc}, the bias of the \gls{LRM} is derived.
In \secref{sec:nparam:simulations} the performance of the different modeling methods are compared by means of simulations.
In \secref{sec:nparam:orderSelection}, some options for model order selection are discussed.
In \secref{sec:nonparametric:truncation}, an approach is developed to further reduce the variance of the \gls{LPM} estimate by truncation of the impulse response.
Finally, conclusions are presented in \secref{sec:conclusion}.

@@

\begin{remark}
This system can be solved reliably only if the columns in \eqref{eq:nparam:designMatrix} are linearly independent.
In practice, this corresponds to the requirements on the input spectrum as stated in \citet{Schoukens2009LPM}: the input spectrum should be sufficiently `rough' to separate the transient contribution from the contribution $\LocalModel{G} \LocalModel{U}$.
This can be obtained e.g. using random phase multisines or random input signals.
\end{remark}

@@

\section{Simulations}
\label{sec:nparam:simulations}

@@

\paragraph{Good signal-to-noise ratios}
The next situation we consider is when $\SNR=60\unit{dB}$ as shown in \figref{fig:nparam:comparison:hiSNR}.
Here, the \gls{LPM} remains severely biased for small datasets ($\nBW < 12$).
However, for sufficiently long experiments ($\nBW > 32$), the bias becomes negligible.
For the \gls{LRM} and \gls{LRIC}, the bias is always much smaller than the variance.
In contrast to \gls{LPM}, the \gls{LRM} already reaches a constant variance level when $\nBW > 10$.
It should also be noted that for the \gls{LRIC}, the variance is almost $10\unit{dB}$ more than the corresponding \gls{LPM} method.
It is hence not advantageous to use the \gls{LRIC} as-is.
For all methods (except the \gls{ETFE} and \gls{LRIC}), the simulations show that the \gls{RMSE} reaches the noise floor minus $\Delta \SNR$.

@@

\section{Conclusions}
\label{sec:conclusion}
In this chapter, the performance of non-parametric \gls{LPM} and \gls{LRM} estimators were compared to observe the \gls{FRF} of resonant systems.
Since the \gls{LRM} is a biased estimator and the \gls{LPM} is asymptotically unbiased, it is important to compare their behavior.
Hence, expressions for the bias of the \gls{LRM} were derived.
For good \glspl{SNR} in resonant systems, it was seen that the \gls{LRM} is always superior to the \gls{LPM}.
For extremely poor \gls{SNR}, i.e. $\SNR \approx 10\unit{dB}$, the \gls{LPM} may perform slightly better than the \gls{LRM} when short measurement records are available.
In this chapter, the \gls{LRIC} was introduced.
This is an extension of the \gls{LRM} which solves the non-linear optimization problem and should thus be able to avoid a bias altogether.
However, it is seen that this estimator is extremely noisy but seems to provide a smaller bias than the \gls{LRM}.
Since the overall \gls{RMS} error for the \gls{LRIC} is much larger than most other methods, the \gls{LRIC} cannot be used in its current form.
